Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-10-13 paper/main.tex
presheaf diagram and nlab citations

 \begin{definition}
$f$ is \textbf{equivariant} when $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}.
 \end{definition}
 \begin{definition}
 $f$ is \textbf{homomorphic} when  $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}. 
 \end{definition}
For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group. For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues. We evaluate whether the encoding is homomorphic when the visual encoding does not have a group structure. For example, partial orders are not invertible; therefore a visual encoding of partially ordered data is structure preserving when the ordering is mapped to equivalently ordered visual encodings.

@@

\begin{definition}\cite{bradleyWhatFunctorDefinitions,bradleyTopologyCategoricalApproach2020} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$, which means it is a function between objects $F: \textbf{ob}(\mathcal{C}) \mapsto \textbf{ob}(\mathcal{D})$ and that for every morphism $f \in Hom(C_1, C_2)$  there is a corresponding function $F: Hom(C1, C2) \mapsto Hom(F(C_1), F( C_2))$.
A \textbf{functor} must satisfy the properties 
\begin{itemize}
  \item \textit{identity}: $F(id_{C}(C)) = id_{D}(F(C))$
  \item \textit{composition}: $F(g)\circ F(f) = F(g\circ f)$ for any composable morphisms $C_{1}\xrightarrow{f} C_2$, $C_2 \xrightarrow{g} C_3$ 
\end{itemize}
$F(C) \in \textbf{ob}(\mathcal{D})$ denotes the object to which an object $C$ is mapped, and $F(f) \in Hom(F_(C_1), F_(C_2))$ denotes the morphism that $f$ is mapped to. 
\end{definition}
Modeling the data container as a functor allows us state that, just like a functor, the container is a map between index space objects and sets of data records that preserve morphisms between index space objects and data records. 
\begin{equation}
  \label{eq:atct:sheaf:functor}
  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
A common way of encapsulating a map from a topological space to a category of sets is as a presheaf
\begin{definition}
  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf, spanier1989algebraic}. 
\end{definition}
Contravariance means that the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects. The presheaf is contravariant because the inclusion morphisms between input objects 
\begin{equation*}
  \label{eq:atct:sheaf:inclusion}
  \iota: \opensetc_1 \rightarrow \opensetc_2
\end{equation*}
are defined such that they correspond to the partial ordering $\openset_1 \subseteq \openset_2$, but the restriction morphisms $\iota^*$ between the sets of sections 
\begin{equation*}
  \label{eq:atct:sheaf:restriction}
  \iota^*: \cgamma{\opensetc_2}{\dtotalc_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc_{\openset_c1}}
\end{equation*}
go in the opposite direction. This is because any arbitrary function defined on an openset $\dsection\restriction_{\openset_2}$ must be definable on a subset of that space $\dsection\restriction_{\openset_1}$. The expectation that data containers implement inclusion and restriction morphisms means that, for example, a data container must ensure that continuous sections can be interpolated.

While presheaves preserve the rules for ensuring continuity in the indexing space and restriction in the sets of data records, sheaves add on conditions for verifying that sections over subspaces are part of a cohesive section over the whole space. 
\begin{definition}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
\begin{itemize}
  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|_{\openset} =  \dsection^{b}|_{\openset}$.
  \item \textit{gluing} the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$
  \end{itemize}
\end{definition}
The gluing axiom says that a distributed representation of a dataset, which is a set of local sections, is equivalent to a section over the union of the opensets of the local sections. The locality axiom asserts that the glued section function is equivalent to a function over the union if they evaluate to the same values. The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local section. Generally, the sheaf asserts the expectation that the data container is implemented such that the connectivity between the opensets (underlying data continuity) is preserved.

@@

\subsection{Data and Graphic Correspondence}
\note{index correspondence}
\label{sec:atct:xi}
There is an expectation that for a visualization to be readable, the visual elements must correspond to distinct data elements\cite{ziemkiewiczEmbeddingInformationVisualization2009} and we can use the properties of sheaves to formally express this correspondence. We first describe the relationship between the graphic indexing space \gbase\ and the data indexing space \dbase\, which we propose is one where multiple graphic indexes map to one data index, and every index in the graphic space can be mapped to an index in the data space. We encode these expectations as the \textcolor{functor}{map} \vindexc, which we define to be a surjective continuous map
\begin{equation}
  \label{eq:atct:xi}
  \vindexc: \opensetgc \textcolor{functor}{\rightarrow} \opensetc 
\end{equation}
between a graphic subspace $\opensetg \subseteq \gbase$ and data subspace $\openset \subseteq \dbase$. The functor $\vindex$ is surjective such that for every point $\dbasepoint \in \openset$ there is a corresponding set of points $\{\gbasepoint | \gbasepoint \in  \vindexpre(k)\}$ for all $\gbasepoint \in \opensetg$. We construct the map as going from graphic to data because that encodes the notion that every visual element traces back to the data in some way. We define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primatives, discussed in \autoref{eq:artist:operator}, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).

\note{add little figure for xi showing it mapping from graphic index space back into data index space}

\subsubsection{Graphic Corresponding to Data}
Since we have defined a function \vindex\ between two spaces $\dbase, \gbase$, we can construct functors that transport sheaves over each space to the other\cite{harder2008lectures}. This allows us to describe what data we expect at each graphic index location and what graphic is expected at each data index location. Transport functors compose the indexing map \vindex\ with the sheave map to say that a record \dsection\ at \dbasepoint\ is at all corrresponding \gbasepoint\ and that a function \gsection\ over one point \gbasepoint\ is the same function at all points $\gbasepoint \in \gbase$ that correspond to the same record index \dbasepoint.

The pushforward (direct image) sheaf establishes which graphic generating function $\gsection$ corresponds to a point $\dbasepoint \in dbase$ in the data base space.   
\begin{definition} Given a sheaf $\sheaf_{\gbase, \gtotal}$ on $\gbase$, the \textbf{pushforward} sheaf  $\vindexpush\sheaf_{\gbase, \gtotal}$ on $\dbase$ is defined as 
  \begin{equation*}
    \vindexpushc(\sheafc_{\gbasec, \gtotalc})(\opensetc)  = \sheafc_{\gbasec, \gtotalc}(\vindexc^{-1}(\opensetc))
  \end{equation*}
for all opensets $\opensetc \subset \dbasec$\cite{harder2008lectures}.
\end{definition}
The pushforward sheaf returns the set of graphic sections over the data base space that corresponds to the graphic space $\vindex^{-1}(\openset) = \opensetgc$. The pushforward functor $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
 \begin{equation}  
  \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}  \ni \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc} 
\end{equation}
such that it provides a way to look up which graphic corresponds with a data index
\begin{equation}
  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)}
\end{equation}
such that $\vindexpush\gsection(\dbasepoint))(\gbasepoint) = \gsection(\gbasepoint)$ for all $\gbasepoint \in \vindexpre(\dbasepoint)$. Therefore, the continuous map $\vindex$ and transport functors $\vindexpull, \vindexpush$ allow us to express the correspondence between graphic section and data section.

\subsubsection{ Data Corresponding to Graphic}
The pullback (inverse image) sheaf establishes which data record  returned by $\dsection$ corresponds to a point $\gbasepoint \in \gbase$ in the graphic base space. 
\begin{definition} \cite{harder2008lectures} Given a sheaf $\sheaf_{\dbase, \dtotal}$ on $\dbase$, the \textbf{pullback} sheaf $\vindexpullc\sheafc_{\dbasec, \dtotalc}$ on $\gbasec$ is defined as the sheaf associated to the presheaf
  \begin{equation*}
    \vindexpullc(\sheafc_{\dbasec, \dtotalc})(\opensetgc) = \sheafc_{\dbasec, \dtotalc}(\vindexc(\opensetgc))   
  \end{equation*}
for $\vindex(\opensetgc) \in \dbase$.
\end{definition}
The pullback sheaf returns the set of data sections over the graphic base space that corresponds to the graphic space $\vindex(\opensetg) = \openset$. The pullback $\vindexpullc$ transports sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
\begin{equation}
  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \ni \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc \dtotalc\restriction_{\opensetgc}
\end{equation}
such that there is a way to then look up what data values correspond with a graphic index
\begin{equation}
  \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindex(\gbasepointc)) = \dsectionc(\dbasepointc)
\end{equation}
As \vindex\ is surjective, there are many points $\gbasepoint \in \opensetg\subseteq\gbase$ in the graphic space that correspond to a single point $\vindex(\gbasepoint) = \dbasepoint$.

\subsubsection{Example: Graphic and Data}
\begin{figure}[h!]
  \includegraphics*[width=1\columnwidth]{xi_scatter.png}
  \caption{The functors $\vindex, \vindexpull, \vindexpush$ establish the correspondence between a data section \dsection\ and a graphic generation function \gsection. For every data record $\dsection(\dbasepoint)$, there is a graphic generating function $\vindexpush\gsection = \gsection|_{\vindexpre\dbasepoint}$. This curried function is here represented as the visual parameters associated with that function to demonstrate that it generates the visual element for that data record. Every data section can also be copied over to the graphic space such that each piece of the (idealized in infinite resolution) graphic $\gsection(\dbasepoint)$ has a corresponding data value $\vindexpull\dsection(\dbasepoint)$.
  \label{fig:atct:morphisms:sheaf}} \note{possibly reorient and expand this diagram to illustrate eq 55}
\end{figure}

Functors between sheaves are a way of expressing the bookkeeping involved in keeping track of which graphic section \gsection\ corresponds to which data section \dsection. The $(\dbasepoint_i, \gbase_i)$ pairing expressed in \autoref{eq:atct:xi} establishes that there is a correspondence between sections evaluated over $\dbasepoint_i$ and $\gbase_i$. This allows us to construct graphic specifications for each data index $\vindexpush\gsection$ and retrieve the data $\vindexpull\dsection$ for any graphic section generating any piece of a graphic. In \autoref{fig:atct:morphisms:sheaf}, the sections return rows in the table from \autoref{fig:related-work:continuity:ktypes}. Each section $\dsection(\dbasepoint_{i})$ has a corresponding visual element produced by evaluating the graphic section on the graphic base space $\gsection|_{\gbase_{i}}$. In \autoref{fig:atct:morphisms:sheaf} the output of $\gsection$ as pixels in an idealized prerendered space is approximated as the circles. Each graphic section $\gsection(\gbasepoint), \gbasepoint \in \gbase_{i}$ that is a piece of the graphic has a corresponding pulled back data value $\vindexpull\dsectionc(\gbasepoint)$. The pullback copies the data points such that $\vindexpull\dsectionc|_{\gbase_i} = \dsection(\dbasepoint_i)$. Each visual element $\gsection|_{\gbase_i}$ is generated from a single record $\dsection(\dbasepoint_i)$; therefore the pushforward is the curried graphic section $\gsection|_{\vindexpre(\dbasepoint_i)}$. The graphic parameterized in \autoref{fig:atct:morphisms:sheaf} is intended as an approximation of $\vindexpush\gsection$.

Visualization specification languages such as vega \cite{heerDeclarative2010} and svg \cite{quintScalable2003} assume that there is a graphic  specification for every point. Interactive tooltips assume that there is a map between the graphics rendered on screen and the data such that a glyph maps back to a record. The functors $(\vindex, \vindexpush, \vindexpull)$ codify the expectation of these mappings between data and graphic space.

@@

\subsection{Map Between Graphics and Data}
In \autoref{sec:atct:sheaves}, the functors $\vindex, \vindexpull, \vindexpushc$ are introduced to describe the expected relationship between screen and data base spaces. In the construction of the artist, the functor $\vindex$ is defined such that the data base space \dbase\ is a deformation retraction\cite{nlab:deformation_retract} of the graphic space $\gbase$. This means that there is a continuous surjective mapping from every point $\dbasepoint\in\dbase$ to a point $\dbasepoint \in dbase$. To simplify matters, in this paper, we construct the graphic space as a constant multiple of the base space such that 
\begin{equation}
  \underbrace{\dbasec\times[0,1]^{n}}_{\gbasec} \textcolor{functor}{\xmapsto{\hspace{1em}\vindexc\hspace{1em}}} \dbasec
\end{equation}
where n is a thickening of the graphic base space $\gbase$ to account for the dimensionality of the output space
\begin{equation*}
  n = \begin{cases}
    dim(\gbase) - dim(\dbase) & dim(\dbase)<dim(\gbase)\\
  0 & otherwise
  \end{cases}
\end{equation*}
because the data dimensionality $\dbase$ may be too small for a graphic representation.

>>>>

 \begin{definition}
$f$ is \textbf{equivariant} when $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant, pittsNominalSetsNames2013}.
 \end{definition}
 \begin{definition}
 $f$ is \textbf{homomorphic} when  $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}. 
 \end{definition}
For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group. For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues. We evaluate whether the encoding is homomorphic when the visual encoding does not have a group structure. For example, partial orders are not invertible; therefore a visual encoding of partially ordered data is structure preserving when the ordering is mapped to equivalently ordered visual encodings.

@@

\begin{definition}\cite{bradleyWhatFunctorDefinitions,bradleyTopologyCategoricalApproach2020} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$, which means it is a function between objects $F: \textbf{ob}(\mathcal{C}) \mapsto \textbf{ob}(\mathcal{D})$ and that for every morphism $f \in Hom(C_1, C_2)$  there is a corresponding function $F: Hom(C1, C2) \mapsto Hom(F(C_1), F( C_2))$.
A \textbf{functor} must satisfy the properties 
\begin{itemize}
  \item \textit{identity}: $F(id_{C}(C)) = id_{D}(F(C))$
  \item \textit{composition}: $F(g)\circ F(f) = F(g\circ f)$ for any composable morphisms $C_{1}\xrightarrow{f} C_2$, $C_2 \xrightarrow{g} C_3$ 
\end{itemize}
$F(C) \in \textbf{ob}(\mathcal{D})$ denotes the object to which an object $C$ is mapped, and $F(f) \in Hom(F_(C_1), F_(C_2))$ denotes the morphism that $f$ is mapped to. 
\end{definition}
Modeling the data container as a functor allows us state that, just like a functor, the container is a map between index space objects and sets of data records that preserve morphisms between index space objects and data records. 
\begin{equation}
  \label{eq:atct:sheaf:functor}
  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
A common way of encapsulating a map from a topological space to a category of sets is as a presheaf
\begin{definition}
  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf, spanier1989algebraic}. 
\end{definition}
Contravariance means that the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects. The presheaf is contravariant because the inclusion morphisms between input objects 
\begin{equation*}
  \label{eq:atct:sheaf:inclusion}
  \iota: \opensetc_1 \rightarrow \opensetc_2
\end{equation*}
are defined such that they correspond to the partial ordering $\openset_1 \subseteq \openset_2$, but the restriction morphisms $\iota^*$ between the sets of sections 
\begin{equation*}
  \label{eq:atct:sheaf:restriction}
  \iota^*: \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}
\end{equation*}
restricts the larger set to the smaller one such that all functions that are continuous over a space must be continuous over a subspace $\Gamma_2 \subseteq \Gamma_1$, where $\Gamma_{i}\coloneqq\Gamma(\openset_{i}, \dtotal\restriction_{\openset_{i}})$.

For example, lets define presheaves $\sheaf_1, \sheaf_2$. These are maps from intervals $\openset_1, \openset_2$ to a set of functions $\Gamma_1, \Gamma_2$ that are continuous over that interval: 
\begin{equation*}
\begin{tikzcd}
  \textcolor{set}{\Gamma_2 = \textcolor{set}{\left\{\begin{array}{c} 
      \textcolor{section}{constant}\\ 
      \textcolor{section}{sin}\\
      \textcolor{section}{cos}\\
    \end{array}\right\}}}               
  \arrow[rr, "\iota^*", color=set] &  & 
  \textcolor{set}{\Gamma_1 = \left\{\begin{array}{c} 
    \textcolor{section}{constant}\\ 
    \textcolor{section}{sin}\\
    \textcolor{section}{cos}\\
    \textcolor{section}{tan}\\
  \end{array}\right\}} \\
  \textcolor{base}{\openset_2 = \left(0,1\right)} 
  \arrow[u, "\sheafc_{1}", maps to, color=sheaf, shift right=4.5]
  &  & 
  \textcolor{base}{\openset_1=\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)} 
  \arrow[ll, "\iota"', color=base] 
  \arrow[u, "\sheafc_{2}"', maps to, color=sheaf, shift right=4.5]
  \end{tikzcd}
\end{equation*}
The constraints of a presheaf functor are that since the $constant, sin, cos$ functions are defined over the interval $\left[0,1\right]$, these functions must also be continuous over the sub-interval $\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)$; therefore the sections in $\Gamma_{2}$ must also be included in the set of sections over the subspace $\Gamma_{1}$. The generalization of this constraint is that data structures that contain continuous functions must support interpolating them over arbitrarily small subspaces.

While presheaves preserve the rules for sets of sections, sheaves add on conditions for gluing individual sections over subspaces into cohesive sections over the whole space. 
\begin{definition}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
\begin{itemize}
  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|_{\openset} =  \dsection^{b}|_{\openset}$.
  \item \textit{gluing} the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$
  \end{itemize}
\end{definition}
The gluing axiom says that a distributed representation of a dataset, which is a set of local sections, is equivalent to a section over the union of the opensets of the local sections. The locality axiom asserts that the glued section function is equivalent to a function over the union if they evaluate to the same values. The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local sections. Generally, the sheaf asserts the expectation that the data container is implemented such that the connectivity between the opensets (indexing subspaces) is preserved.

@@

\subsection{Data Index and Graphic Index Correspondence}
\label{sec:atct:xi}
There is an expectation that for a visualization to be readable, the visual elements must correspond to distinct data elements\cite{ziemkiewiczEmbeddingInformationVisualization2009} and we can use the properties of sheaves to formally express this correspondence. We first describe the relationship between the graphic indexing space \gbase\ and the data indexing space \dbase\, which we propose is one where multiple graphic indexes map to one data index, and every index in the graphic space can be mapped to an index in the data space. We encode these expectations as the \textcolor{functor}{map} \vindexc, which we define to be a surjective continuous map
\begin{equation}
  \label{eq:atct:xi}
  \vindexc: \opensetgc \textcolor{functor}{\rightarrow} \opensetc 
\end{equation}
between a graphic subspace $\opensetg \subseteq \gbase$ and data subspace $\openset \subseteq \dbase$. The functor $\vindex$ is surjective such that for every point $\dbasepoint \in \openset$ there is a corresponding set of points $\{\gbasepoint | \gbasepoint \in  \vindexpre(k)\}$ for all $\gbasepoint \in \opensetg$.

We construct the map as going from graphic to data because that encodes the notion that every visual element traces back to the data in some way. As exemplified in \autoref{fig:atct:morphisms:sheaf}, we define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primatives, discussed in \autoref{eq:artist:operator}, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).

\subsubsection{Data and Graphic Correspondence}
Since we have defined a function \vindex\ between two spaces $\dbase, \gbase$, we can then construct functors that transport sheaves over each space to the other\cite{harder2008lectures}. This allows us to describe what data we expect at each graphic index location and what graphic is expected at each data index location. Transport functors compose the indexing map \vindex\ with the sheave map to say that a record \dsection\ at \dbasepoint\ is at all corrresponding \gbasepoint\ and that a function \gsection\ over one point \gbasepoint\ is the same function at all points $\gbasepoint \in \gbase$ that correspond to the same record index \dbasepoint.

\paragraph{\textbf{Graphic Corresponding to Data}}
The pushforward (direct image) sheaf establishes which graphic generating function $\gsection$ corresponds to a point $\dbasepoint \in dbase$ in the data base space.   
\begin{definition} Given a sheaf $\sheaf_{\gbase, \gtotal}$ on $\gbase$, the \textbf{pushforward} sheaf  $\vindexpush\sheaf_{\gbase, \gtotal}$ on $\dbase$ is defined as 
  \begin{equation*}
    \vindexpushc(\sheafc_{\gbasec, \gtotalc})(\opensetc)  = \sheafc_{\gbasec, \gtotalc}(\vindexc^{-1}(\opensetc))
  \end{equation*}
for all opensets $\opensetc \subset \dbasec$\cite{harder2008lectures}.
\end{definition}
The pushforward sheaf returns the set of graphic sections over the data base space that corresponds to the graphic space $\vindex^{-1}(\openset) = \opensetgc$. The pushforward functor $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
 \begin{equation}  
  \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}  \ni \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc} 
\end{equation}
such that it provides a way to look up which graphic corresponds with a data index
\begin{equation}
  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)}
\end{equation}
such that $\vindexpush\gsection(\dbasepoint))(\gbasepoint) = \gsection(\gbasepoint)$ for all $\gbasepoint \in \vindexpre(\dbasepoint)$. Therefore, the continuous map $\vindex$ and transport functors $\vindexpull, \vindexpush$ allow us to express the correspondence between graphic section and data section.

\paragraph{\textbf{Data Corresponding to Graphic}}
The pullback (inverse image) sheaf establishes which data record  returned by $\dsection$ corresponds to a point $\gbasepoint \in \gbase$ in the graphic base space. 
\begin{definition} \cite{harder2008lectures} Given a sheaf $\sheaf_{\dbase, \dtotal}$ on $\dbase$, the \textbf{pullback} sheaf $\vindexpullc\sheafc_{\dbasec, \dtotalc}$ on $\gbasec$ is defined as the sheaf associated to the presheaf
  \begin{equation*}
    \vindexpullc(\sheafc_{\dbasec, \dtotalc})(\opensetgc) = \sheafc_{\dbasec, \dtotalc}(\vindexc(\opensetgc))   
  \end{equation*}
for $\vindex(\opensetgc) \in \dbase$.
\end{definition}
The pullback sheaf returns the set of data sections over the graphic base space that corresponds to the graphic space $\vindex(\opensetg) = \openset$. The pullback $\vindexpullc$ transports sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
\begin{equation}
  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \ni \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc \dtotalc\restriction_{\opensetgc}
\end{equation}
such that there is a way to then look up what data values correspond with a graphic index
\begin{equation}
  \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindex(\gbasepointc)) = \dsectionc(\dbasepointc)
\end{equation}
As \vindex\ is surjective, there are many points $\gbasepoint \in \opensetg\subseteq\gbase$ in the graphic space that correspond to a single point $\vindex(\gbasepoint) = \dbasepoint$.

\subsubsection{Example: Graphic and Data}
\begin{figure}[h!]
  \includegraphics*[width=1\columnwidth]{xi_sin.png}
  \caption{The data consists of the $sin$ and $cos$ functions over a unit circle base space. We choose to visualize this as a circle and two line plots. The indexing function \vindex\ bookkeeps which parts of the circle and each curve correspond to each point on the unit circle. The pushforward $\vindexpush$ matches each point in the data space to the specification of the graphic at that point, while the pullback $\vindexpull$ matches each point in the graphic space to the data over that point. \label{fig:atct:morphisms:sheaf}} 
\end{figure}

Functors between sheaves are a way of expressing the bookkeeping involved in keeping track of which graphic section \gsection\ corresponds to which data section \dsection. The $(\dbasepoint_i, \gbase_i)$ pairing expressed in \autoref{eq:atct:xi} establishes that there is a correspondence between sections evaluated over $\dbasepoint_i$ and $\gbase_i$. This allows us to construct graphic specifications for each data index $\vindexpush\gsection$ and retrieve the data $\vindexpull\dsection$ for any graphic section generating any piece of a graphic. In \autoref{fig:atct:morphisms:sheaf}, the visualization is a graphic representation of a unit circle and the sin and cosine curves on that interval. The index lookup $\vindex$ describes which parts of the circle and curves are generated from which points on the unit circle. Given this correspondance, the pullback $\vindexpull$ looks up which values are being represented in a given part of the graphic. The pushforward $\vindexpush$ describes how a graphic is supposed to look for each point in the data space. The graphic parameterized in \autoref{fig:atct:morphisms:sheaf} is intended as an approximation of $\vindexpush\gsection$.

Visualization specification languages such as vega \cite{heerDeclarative2010} and svg \cite{quintScalable2003} assume that there is a graphic  specification for every point. Interactive tooltips assume that there is a map between the graphics rendered on screen and the data such that a glyph maps back to a record. The expectation that these mappings between data and graphic space exist are encapsulated in the functors $(\vindex, \vindexpush, \vindexpull)$.

@@

\subsection{Map Between Graphics and Data}
In \autoref{sec:atct:sheaves}, the functors $\vindex, \vindexpull, \vindexpushc$ are introduced to describe the expected relationship between screen and data base spaces. In the construction of the artist, the functor $\vindex$ is defined such that the data base space \dbase\ is a deformation retraction\cite{nlab:deformation_retract,hatcherAlgebraicTopology2002} of the graphic space $\gbase$. This means that there is a continuous surjective mapping from every point $\dbasepoint\in\dbase$ to a point $\dbasepoint \in dbase$. To simplify matters, in this paper, we construct the graphic space as a constant multiple of the base space such that 
\begin{equation}
  \underbrace{\dbasec\times[0,1]^{n}}_{\gbasec} \textcolor{functor}{\xmapsto{\hspace{1em}\vindexc\hspace{1em}}} \dbasec
\end{equation}
where n is a thickening of the graphic base space $\gbase$ to account for the dimensionality of the output space
\begin{equation*}
  n = \begin{cases}
    dim(\gbase) - dim(\dbase) & dim(\dbase)<dim(\gbase)\\
  0 & otherwise
  \end{cases}
\end{equation*}
because the data dimensionality $\dbase$ may be too small for a graphic representation.
